The Right Way to Search Evolving Graphs

### 2016-01-28 paper.tex
change abstract

\begin{abstract}
We generalize the breadth first search (BFS) algorithm to evolving graphs and
prove its correctness.
The BFS generates paths with nontrivial structure, which we term temporal paths.
We also show with simple counterexamples that products of adjacency matrices alone
do not have the correct combinatorics of enumerating temporal paths.
Generalizing BFS to evolving graphs requires careful tracking of active nodes
which participate in the definition of edges at specific times.
We discuss how the structure of temporal paths is reflected in adjacency matrix
representations of evolving graphs,
as well as implications for real world applications such as the analysis of citation networks.
\end{abstract}

@@

In this paper, we explain how the BFS algorithm can be applied correctly on
evolving graphs. We also provide a counterexample demonstrating that some na\"ive
analyses in terms of discrete path sums over the time-independent adjacency
matrices lead to incorrect results. The correct generalization to evolving graphs
is formally equivalent to BFS over a particular static graph generated by connecting
together active notes of the evolving graph. We also show that BFS over evolving
graphs can be interpreted algebraically as computing a sequence of vectors formed
by repeatedly multiplying a starting vector with the adjacency matrix of the
equivalent static graph, and in so doing reveals interesting connections between
properties of the BFS algorithm and properties of the adjacency matrix. Finally,
we explain how BFS on evolving graphs may be applied to study dynamical processes
over citation networks.

>>>>

\begin{abstract}
We present a correctly verified breadth first search (BFS) algorithm for evolving graphs. Using simple examples, we show na\"ive matrix-matrix multiplication on time-dependent adjacency matrices miscounts the number of temporal paths and therefore should be avoided if temporal path is a component of the algorithm. We formulate our
evolving graph BFS using the language of linear algebra and present an adjacency matrix representation of the equivalent static graph, which is used to study
interesting properties about the original evolving graphs. 
% We generalize the breadth first search (BFS) algorithm to evolving graphs and
% prove its correctness.
% The BFS generates paths with nontrivial structure, which we term temporal paths.
% We also show with simple counterexamples that products of adjacency matrices alone
% do not have the correct combinatorics of enumerating temporal paths.
% Generalizing BFS to evolving graphs requires careful tracking of active nodes
% which participate in the definition of edges at specific times.
% We discuss how the structure of temporal paths is reflected in adjacency matrix
% representations of evolving graphs,
% as well as implications for real world applications such as the analysis of citation networks.
\end{abstract}

@@

In this paper, we explain how the BFS algorithm can be applied correctly on
evolving graphs. We also provide a counterexample demonstrating that some na\"ive
analyses in terms of discrete path sums over the time-dependent adjacency
matrices lead to incorrect results. The correct generalization to evolving graphs
is formally equivalent to BFS over a particular static graph generated by connecting
together active notes of the evolving graph. We also show that BFS over evolving
graphs can be interpreted algebraically as computing a sequence of vectors formed
by repeatedly multiplying a starting vector with the adjacency matrix of the
equivalent static graph, and in so doing reveals interesting connections between
properties of the BFS algorithm and properties of the adjacency matrix. Finally,
we explain how BFS on evolving graphs may be applied to study dynamical processes
over citation networks.
